Hazelnut Live: Type-Theoretic Foundations for Running Incomplete Programs

### 2017-06-09 calculus.tex
fixed cast syntax, cast stepping rule
added missing requirement for substitution typing

\begin{figure}[h]
$\arraycolsep=4pt\begin{array}{rllllll}
\mathsf{HTyp} & \htau & ::= &
  b ~\vert~
  \tarr{\htau}{\htau} ~\vert~
  % \tprod{\htau}{\htau} ~\vert~
  % \tsum{\htau}{\htau} ~\vert~
  \tehole\\
\mathsf{HExp} & \hexp & ::= &
  c ~\vert~
  x ~\vert~
  \halam{x}{\htau}{\hexp} ~\vert~
  \hap{\hexp}{\hexp} ~\vert~
  % \hpair{\hexp}{\hexp} ~\vert~
  % \hprj{i}{\hexp} ~\vert~
  % \hinj{i}{\hexp} ~\vert~
  % \hcase{\hexp}{x}{\hexp}{x}{\hexp} ~\vert~
  % \hadd{\hexp}{\hexp} ~\vert~
  \hehole{u} ~\vert~
  \hhole{\hexp}{u} ~\vert~
  {\hlam{x}{\hexp}} ~\vert~
  \hexp : \htau\\
 \mathsf{DHExp} & \dexp  & ::= &
  c ~\vert~
  x ~\vert~
  {\halam{x}{\htau}{\dexp}} ~\vert~
  \hap{\dexp}{\dexp} ~\vert~
  % \hpair{\dexp}{\dexp} ~\vert~
  % \hprj{i}{\dexp} ~\vert~
  % \hinj{i}{\dexp} ~\vert~
  % \hcase{\dexp}{x}{\dexp}{x}{\dexp} ~\vert~
  % \hadd{\dexp}{\dexp} ~\vert~
  \dehole{\mvar}{\subst} ~\vert~
  \dhole{\dexp}{\mvar}{\subst} ~\vert~
  \dcast{\dexp}{\htau}\\
\end{array}$
\caption{Syntax of H-types, H-expressions and dynamic H-expressions. We write $x$ to range over variables, $u$ over hole names and $\sigma$ over finite substitutions.}
\label{fig:HTyp}
\label{fig:HExp}
\end{figure}

@@

\inferrule[ESAp1]{
  \hsyn{\hGamma}{\hexp_1}{\tehole}\\
  \expandAna{\hGamma}{\hexp_2}{\tehole}{\dexp_2}{\htau_2}{\Delta_2}\\
  \expandAna{\hGamma}{\hexp_1}{\tarr{\htau_2}{\tehole}}{\dexp_1}{\htau_1}{\Delta_1}
}{
  \expandSyn{\hGamma}{\hap{\hexp_1}{\hexp_2}}{\tehole}{\hap{(\dcast{\dexp_1}{\tarr{\htau_2}{\tehole}})}{\dexp_2}}{\Dunion{\Delta_1}{\Delta_2}}
}

\inferrule[ESAp2]{
  \expandSyn{\hGamma}{\hexp_1}{\tarr{\htau_2}{\htau}}{\dexp_1}{\Delta_1}\\
  \expandAna{\hGamma}{\hexp_2}{\htau_2}{\dexp_2}{\htau'_2}{\Delta_2}\\
  \htau_2 \neq \htau'_2
}{
  \expandSyn{\hGamma}{\hap{\hexp_1}{\hexp_2}}{\htau}{\hap{\dexp_1}{\dcast{\dexp_2}{\htau_2}}}{\Dunion{\Delta_1}{\Delta_2}}
}

@@

\inferrule[ESNEHole]{
  \expandSyn{\hGamma}{\hexp}{\htau}{\dexp}{\Delta}
}{
  \expandSyn{\hGamma}{\hhole{\hexp}{u}}{\tehole}{\dhole{\dexp}{u}{\idof{\hGamma}}}{\Delta, \Dbinding{u}{\hGamma}{\tehole}}
}\\
%
\inferrule[ESAsc1]{
  \expandAna{\hGamma}{\hexp}{\htau}{\dexp}{\htau'}{\Delta}\\
  \htau \neq \htau'
}{
  \expandSyn{\hGamma}{\hexp : \htau}{\htau}{\dcast{\dexp}{\htau}}{\Delta}
}

@@

\inferrule[TACast]{
  \hasType{\Delta}{\Gamma}{\dexp}{\htau'}\\
  \tconsistent{\htau}{\htau'}
}{
  \hasType{\Delta}{\hGamma}{\dcast{\dexp}{\htau}}{\htau}
}
\end{mathpar}
\caption{Type Assignment for Dynamic H-Expressions}
\label{fig:hasType}
\end{figure}

\begin{figure}[!ht]
  \begin{definition}
    $\hasType{\Delta}{\hGamma}{\sigma}{\hGamma'}$ iff for each $\dexp/x \in \sigma$, we have $x : \htau \in \hGamma'$ and $\hasType{\Delta}{\hGamma}{\dexp}{\htau}$.
  \end{definition}
  \caption{substitution type assignment}
\end{figure}

@@

\inferrule[ICast]
{\isIndet{\dexp}}
{\isIndet{\dcast{\dexp}{\htau}}}

@@

\inferrule[ERCastError~\todo{??}]
% XXX use macro type assignment judgement here
{\hasType{\Delta}{\todo{\emptyset}}{e}{\htau_2}\\
 \tinconsistent{\htau_1}{\htau_2}}
{\isErr{\Delta}{\dcast{\dexp}{\htau_1}}}

@@

\inferrule[ERCast]
% XXX use macro type assignment judgement here
{\isErr{\Delta}{\dexp}}
{\isErr{\Delta}{\dcast{\dexp}{\htau_1}}}

@@

\judgbox{\stepsTo{\dexp_1}{\dexp_2}}{$\dexp_1$ steps to $\dexp_2$}
\begin{mathpar}
\Infer{STHole}
{\stepsTo{\dexp_1}{\dexp_2} }
{\stepsTo{\dhole{\dexp_1}{\mvar}{\subst}}{\dhole{\dexp_2}{\mvar}{\subst}}}
\and
\Infer{STCast}
% XXX use macro type assignment judgement here
{ \vdash e : \htau_2 \\ \tconsistent{\htau_1}{\htau_2}}
{\stepsTo{\dcast{\dexp}{\htau_1}}{\dexp}}
\\
\Infer{STAp1}
{\stepsTo{\dexp_1}{\dexp_1'}}
{\stepsTo{\dap{\dexp_1}{\dexp_2}}{\dap{\dexp_1'}{\dexp_2}}}
\and
\Infer{STAp2}
{ \isFinal{\dexp_1} \\ \stepsTo{\dexp_2}{\dexp_2'}}
{\stepsTo{\dap{\dexp_1}{\dexp_2}}{\dap{\dexp_1}{\dexp_2'}}}
\and
\Infer{STAp$\beta$}
{ \isFinal{\dexp_2} }
{\stepsTo{\dapP{\dlam{x}{\htau}{\dexp_1}}{\dexp_2}}{ [\dexp_2/x]\dexp_1 }}
\end{mathpar}
\caption{Structural Dynamics}
\label{fig:isValue}
\label{fig:isIndet}
\label{fig:stepsTo}
\end{figure}

>>>>

\begin{figure}[h]
$\arraycolsep=4pt\begin{array}{rllllll}
\mathsf{HTyp} & \htau & ::= &
  b ~\vert~
  \tarr{\htau}{\htau} ~\vert~
  % \tprod{\htau}{\htau} ~\vert~
  % \tsum{\htau}{\htau} ~\vert~
  \tehole\\
\mathsf{HExp} & \hexp & ::= &
  c ~\vert~
  x ~\vert~
  \halam{x}{\htau}{\hexp} ~\vert~
  \hap{\hexp}{\hexp} ~\vert~
  % \hpair{\hexp}{\hexp} ~\vert~
  % \hprj{i}{\hexp} ~\vert~
  % \hinj{i}{\hexp} ~\vert~
  % \hcase{\hexp}{x}{\hexp}{x}{\hexp} ~\vert~
  % \hadd{\hexp}{\hexp} ~\vert~
  \hehole{u} ~\vert~
  \hhole{\hexp}{u} ~\vert~
  {\hlam{x}{\hexp}} ~\vert~
  \hexp : \htau\\
 \mathsf{DHExp} & \dexp  & ::= &
  c ~\vert~
  x ~\vert~
  {\halam{x}{\htau}{\dexp}} ~\vert~
  \hap{\dexp}{\dexp} ~\vert~
  % \hpair{\dexp}{\dexp} ~\vert~
  % \hprj{i}{\dexp} ~\vert~
  % \hinj{i}{\dexp} ~\vert~
  % \hcase{\dexp}{x}{\dexp}{x}{\dexp} ~\vert~
  % \hadd{\dexp}{\dexp} ~\vert~
  \dehole{\mvar}{\subst} ~\vert~
  \dhole{\dexp}{\mvar}{\subst} ~\vert~
  \dcast{\htau}{\dexp}\\
\end{array}$
\caption{Syntax of H-types, H-expressions and dynamic H-expressions. We write $x$ to range over variables, $u$ over hole names and $\sigma$ over finite substitutions.}
\label{fig:HTyp}
\label{fig:HExp}
\end{figure}

@@

\inferrule[ESAp1]{
  \hsyn{\hGamma}{\hexp_1}{\tehole}\\
  \expandAna{\hGamma}{\hexp_2}{\tehole}{\dexp_2}{\htau_2}{\Delta_2}\\
  \expandAna{\hGamma}{\hexp_1}{\tarr{\htau_2}{\tehole}}{\dexp_1}{\htau_1}{\Delta_1}
}{
  \expandSyn{\hGamma}{\hap{\hexp_1}{\hexp_2}}{\tehole}{\hap{(\dcast{\tarr{\htau_2}{\tehole}}{\dexp_1})}{\dexp_2}}{\Dunion{\Delta_1}{\Delta_2}}
}

\inferrule[ESAp2]{
  \expandSyn{\hGamma}{\hexp_1}{\tarr{\htau_2}{\htau}}{\dexp_1}{\Delta_1}\\
  \expandAna{\hGamma}{\hexp_2}{\htau_2}{\dexp_2}{\htau'_2}{\Delta_2}\\
  \htau_2 \neq \htau'_2
}{
  \expandSyn{\hGamma}{\hap{\hexp_1}{\hexp_2}}{\htau}{\hap{\dexp_1}{\dcast{\htau_2}{\dexp_2}}}{\Dunion{\Delta_1}{\Delta_2}}
}

@@

\inferrule[ESNEHole]{
  \expandSyn{\hGamma}{\hexp}{\htau}{\dexp}{\Delta}
}{
  \expandSyn{\hGamma}{\hhole{\hexp}{u}}{\tehole}{\dhole{\dexp}{u}{\idof{\hGamma}}}{\Delta, \Dbinding{u}{\hGamma}{\tehole}}
}\\
%
\inferrule[ESAsc1]{
  \expandAna{\hGamma}{\hexp}{\htau}{\dexp}{\htau'}{\Delta}\\
  \htau \neq \htau'
}{
  \expandSyn{\hGamma}{\hexp : \htau}{\htau}{\dcast{\htau}{\dexp}}{\Delta}
}

@@

\inferrule[TACast]{
  \hasType{\Delta}{\Gamma}{\dexp}{\htau'}\\
  \tconsistent{\htau}{\htau'}
}{
  \hasType{\Delta}{\hGamma}{\dcast{\htau}{\dexp}}{\htau}
}
\end{mathpar}
\caption{Type Assignment for Dynamic H-Expressions}
\label{fig:hasType}
\end{figure}

\begin{figure}[!ht]
  \begin{definition}
    $\hasType{\Delta}{\hGamma}{\sigma}{\hGamma'}$ iff $\domof{\sigma} = \domof{\Delta}$ and for each $\dexp/x \in \sigma$, we have $x : \htau \in \hGamma'$ and $\hasType{\Delta}{\hGamma}{\dexp}{\htau}$.
  \end{definition}
  \caption{substitution type assignment}
\end{figure}

@@

\inferrule[ICast]
{\isIndet{\dexp}}
{\isIndet{\dcast{\htau}{\dexp}}}

@@

\inferrule[ERCastError~\todo{??}]
% XXX use macro type assignment judgement here
{\hasType{\Delta}{\todo{\emptyset}}{e}{\htau_2}\\
 \tinconsistent{\htau_1}{\htau_2}}
{\isErr{\Delta}{\dcast{\htau_1}{\dexp}}}

@@

\inferrule[ERCast]
% XXX use macro type assignment judgement here
{\isErr{\Delta}{\dexp}}
{\isErr{\Delta}{\dcast{\htau_1}{\dexp}}}

@@

\judgbox{\stepsTo{\dexp_1}{\dexp_2}}{$\dexp_1$ steps to $\dexp_2$}
\begin{mathpar}
\Infer{STHole}
{\stepsTo{\dexp_1}{\dexp_2} }
{\stepsTo{\dhole{\dexp_1}{\mvar}{\subst}}{\dhole{\dexp_2}{\mvar}{\subst}}}
\and
\Infer{STCast}
{\hasType{\Delta}{\emptyset}{\dexp}{\htau_2} \\
 \tconsistent{\tau_1}{\tau_2}}
{\stepsTo{\dcast{\htau_1}{\dexp}}{\dexp}}
\\
\Infer{STAp1}
{\stepsTo{\dexp_1}{\dexp_1'}}
{\stepsTo{\dap{\dexp_1}{\dexp_2}}{\dap{\dexp_1'}{\dexp_2}}}
\and
\Infer{STAp2}
{ \isFinal{\dexp_1} \\ \stepsTo{\dexp_2}{\dexp_2'}}
{\stepsTo{\dap{\dexp_1}{\dexp_2}}{\dap{\dexp_1}{\dexp_2'}}}
\and
\Infer{STAp$\beta$}
{ \isFinal{\dexp_2} }
{\stepsTo{\dapP{\dlam{x}{\htau}{\dexp_1}}{\dexp_2}}{ [\dexp_2/x]\dexp_1 }}
\end{mathpar}
\caption{Structural Dynamics}
\label{fig:isValue}
\label{fig:isIndet}
\label{fig:stepsTo}
\end{figure}
